---
title: "p8105_hw2_rw2708"
author: "Ran"
date: "9/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```


#Problem 1
The chunk below serves to read and clean the Mr. Trash Wheel sheet.
```{r Trash_Wheel,message=FALSE, collapse=TRUE}
#read the excel file
wheel_data = read_excel("hw2data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", skip = 1, sheet = 1) %>% 
    #use reasonable variable names
    janitor::clean_names() %>%
  
    #omit rows that do not include dumpster-specific data
    select(dumpster:homes_powered) %>%
    drop_na(dumpster) %>% 
    
    #round the number of sports balls to the nearest integer and converts the result to an integer variable
    mutate(sports_balls = round(sports_balls,digits = 0)) %>% 
    mutate(sports_balls = as.integer(sports_balls)) 

wheel_data

```


The chunk below serves to read and clean the precipitation data for 2017 and 2018.
```{r precipitation2018}
#read the excel file (2018 precipitation data)
precipitation2018_data = read_excel("hw2data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",skip = 1, sheet = 5,range = "A2:B15") %>% 
  
  #use reasonable variable names
  janitor::clean_names() %>%
  
  # omit rows without precipitation data
  drop_na() %>% 
  
  # add a variable year 
  mutate(year = 2018)

precipitation2018_data

#read the excel file (2017 precipitation data)
precipitation2017_data = read_excel("hw2data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",skip = 1, sheet = 6,range = "A2:B14") %>% 
  
   #use reasonable variable names
  janitor::clean_names() %>%
  
  # omit rows without precipitation data
  drop_na() %>% 
  
  # add a variable year 
  mutate(year = 2017)

precipitation2017_data

#combine the above two datasets
combine_data = bind_rows(precipitation2017_data, precipitation2018_data) %>% 
  janitor::clean_names() %>%
  
  #convert month to a character variable
  mutate(month = month.name[month])  

combine_data

```

Description: 1) For the first dataset (wheel_data), there exist `r nrow(wheel_data)` observations, with `r ncol(wheel_data)` variables, and it's key variables are `r colnames(wheel_data)[1:6]`.The first four variables `r colnames(wheel_data)[1:4]` indicate each dumpster at different time, and the following two variables`r colnames(wheel_data)[5:6]` indicate the amount of trash that dumped into each dumpster. 2) For the second dataset (precipitation2018_data), it contains the precipitation data in 2018. There exist `r nrow(precipitation2018_data)` observations, with `r ncol(precipitation2018_data)` variables, and it's key variables are `r ls(precipitation2018_data)`.
3) For the third dataset (precipitation2017_data), it contains the precipitation data in 2017. There exist `r nrow(precipitation2017_data)` observations, with `r ncol(precipitation2017_data)` variables, and it's key variables are `r ls(precipitation2017_data)`. 4) As for the last dataset (combine_data), it contains the precipitation data in 2017 and 2018. There exist `r nrow(combine_data)` observations, with `r ncol(combine_data)` variables, and it's key variables are `r ls(combine_data)`. These three variables indicate the total precipitation in each month from 2017 to 2018. 

Furthermore,the median number of sports balls in a dumpster in 2017 is `r median(pull(wheel_data[pull(wheel_data,year) ==2017,],sports_balls))`. And the total precipitation in 2018 is `r sum(pull(precipitation2018_data,total))`.


#Problem 2
```{r pol_month}
#clean the data in pols-month.csv
pol_data = read_csv(file = "hw2data/pols-month.csv") %>% 
  janitor::clean_names() %>%
 separate(mon, into = c("year","month","day"), sep = "-") %>%
 mutate(month = as.integer(month)) %>% 
 mutate(month = month.name[month]) %>% 
  select(year, month, day, prez_gop,prez_dem, everything()) %>%
    pivot_longer(
    prez_gop:prez_dem,
    names_to = "president", 
    names_prefix = "prez_",
    values_to = "president_values") %>% 
    filter(president_values != "0") %>% 
    select(-day,-president_values)

pol_data

#clean the data in snp.csv using a similar process to the above
snp_data = read_csv(file = "hw2data/snp.csv") %>%
   janitor::clean_names() %>%
  separate(date, into = c("month","day","year"), sep = "/") %>%
  arrange(year,month) %>% 
  mutate(month = as.integer(month)) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day) %>% 
  select(year, month, close)

snp_data

#tidy the unemployment data
unemployment_data = read_csv(file = "hw2data/unemployment.csv") %>%
               pivot_longer(
                 Jan:Dec,
                 names_to = "month",
                 values_to = "unemployment") %>% 
              rename(year = Year) %>% 
       mutate(year = as.character(year)) %>% 
       mutate(month = str_subset(month.name, str_to_sentence(month)))

unemployment_data

#Join the datasets by merging snp into pols, and merging unemployment into the result.
pol_snp = left_join(pol_data,snp_data, by=c("year","month"))
 
final_data = left_join(pol_snp,unemployment_data, by = c("year","month"))

final_data

```

Description: 1) For the first dataset (pol_data), which contains number of national politicians who are democratic or republican, and the infromation about the party that president came from over time. The data has `r nrow(pol_data)` observations, with `r ncol(pol_data)` variables, and it's key variables are `r colnames(pol_data)[1:2]` , and `r colnames(pol_data)[9]`. These variables indicate the party that president came from over time. Furthermore,the range of years is (`r range(pull(pol_data,year))`). 2) For the second dataset (snp_data), which contains figures of S&P close price (stock market information) over time. The data has `r nrow(snp_data)` observations, with `r ncol(snp_data)` variables, and it's key variables are `r ls(snp_data)`. And the range of years is (`r range(pull(snp_data,year))`).  3) For the third dataset (unemployment_data), which contains the percentage of unemployment over time. The data has `r nrow(unemployment_data)` observations, with `r ncol(unemployment_data)` variables, and it's key variables are `r ls(unemployment_data)`. In addition,the range of years is (`r range(pull(unemployment_data,year))`). 4) For the last dataset (final_data), which consist of the above three datasets. It has `r nrow(final_data)` observations, with `r ncol(final_data)` variables, and it's key variables are `r colnames(final_data)[1:2]` and `r colnames(final_data)[9:11]`.These variables indicate the party that president comes from over time, along with the unemployment and stock market information in the same period. Furthermore,the range of years is (`r range(pull(final_data,year))`).


#Problem 3
```{r baby_names}
#Load and tidy the data of popularity of baby names
baby_names = read_csv(file = 'hw2data/Popular_Baby_Names.csv') %>% 
   janitor::clean_names() %>%
mutate(childs_first_name = str_to_sentence(childs_first_name)) %>% 
mutate(ethnicity = str_to_lower(ethnicity)) %>% 
mutate(ethnicity = replace(ethnicity,ethnicity == "white non hisp","white non hispanic")) %>% 
mutate(ethnicity = replace(ethnicity,ethnicity == "asian and paci","asian and pacific")) %>% 
mutate(ethnicity = replace(ethnicity,ethnicity == "black non hisp","black non hispanic")) %>% 
mutate(gender = str_to_lower(gender)) %>% 
distinct()

#Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time
olivia_data = filter(baby_names, childs_first_name == 'Olivia') %>% 
  pivot_wider(id_cols = ethnicity,
              names_from = "year_of_birth",
              values_from = "rank") 

#Produce a similar table showing the most popular name among male children over time
male_child = filter(baby_names, gender == 'male',rank == "1") %>% 
   pivot_wider(id_cols = ethnicity,
              names_from = "year_of_birth",
              values_from = "childs_first_name") 

#Create a data of male, white non-hispanic children born in 2016
male_white_nonhis = filter(baby_names, gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016")

#Produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis)
ggplot(male_white_nonhis,aes(x = rank ,y = count, color = rank)) + geom_point()


```

The above scatter plot shows the  number of male, white non-hispanic children born in 2016 with a name (y axis) against the rank in popularity of that name (x axis).The trend of the scatter plot is that counts decrease from approximately 260 to nearly zero as rank increases from 0 to 100.



